---
title: "Factors that Affect the Credit Limits"
author: "Xinxuan Lin"
date: "March 26, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

A higher credit limit is always a big attraction for people to apply for credit cards. However, for two people who have similar performance on the covariates, they can be issued with different credit limits. In this report, the dataset *BankChurners* provided by S. Goyal were used to determine the association between credit limits, number of transaction, customer's gender, and their income. Based on the result of the linear regression model and the AIC backward method, there is a positive relationship between credit limits and the number of total transactions and income. Moreover, females are possible to get higher credit limits than males.\newline

## 1. Introduction

As the widely use of credit cards, credit card companies have to offer credit limits more carefully as a way to prevent financial loss. However, from the customer's perspective, a higher credit limit is always a big motivation for them to use credit cards. Hence, the analysis of the dataset *BankChurners* could help them understand how credit card companies offer credit limits based on their performances, and what they could do to increase their credit limits.\newline

The credit limit is the maximum amount of what the credit card issuer will allow the client to borrow from them (Porter et al., 2020). The initial credit limits are based on the information provided by the customers in the application form. Based on customers' performances, the credit limits would be adjusted. However, people always wondering why others can have higher credit limits when they have similar performances.\newline

This report would determine the important factors that credit card companies would use when offering credit limits. Customers' age, gender income, the frequency of transactions, and education level were first considered to build a mulivariable linear model. These five explanatory variables were first introduced to the model, because it was the basic and common information that companies would get from customers. After getting this initial model, the AIC backward elimination method was applied to find the actual important factors out of these five covariates. \newline

*Bankchurners* provided by S. Goyal on *Kaggle* was used in this report to determining the association between credit limits and mentioned variables. Information about the dataset was included in the Data section (Section 2.1); the model for prediction of credit limits was introduced in Model section (Section 2.2); Result section (Section 3) explained the relationship between credit limits and important covariates. The limitation of the study was included in Section 4. Conclusion section summarized the association between credit limits and mentioned factors.\newline

## 2.1 Data 

The data *BankChurners* was provided by S. Goyal on Kaggle. It contains personal information of 10,127 customers in the bank. Due to the privacy issue, the names of customers and bank were not provided. Customers were listed out by using 9 digits numbers. Hence, the analysis of this result was based on an observational study.\newline

### 2.1.1 Method of Collecting Data
The data were collected based on the application form that customers submitted, and also further financial performance after they got their credit cards. The performances were recorded in the bank system.\newline

The response of "age", "gender", "income", and "education level" were collected based on the application form that respondent completed. Further updates of the information are respondents' voluntary action. The non-response answer were recorded as "unknown" in the system. \newline  
Frequency of transaction was collected by using the bank monitoring system. Banks have a record of each transaction that customers make by using their bank accounts.\newline

### 2.1.2 Variables
**Education Level**:\newline
 Education level is a categorical variable. It has 7 categories as following:\newline
 1.College \newline
 2.Doctorate \newline
 3.Graduate \newline
 4.High School \newline
 5.Post-Graduate \newline
 6.Uneducated \newline
 7.Unknown \newline
 \newline
 
 **Income_Category:**\newline
The variable income in the dataset is also recorded as category. It is separated into 6 different categories:\newline
1.\$120K + \newline
2.\$40K - \$60K \newline
3.\$60K - \$80K \newline
4.\$80K - \$120K \newline
5. Less than \$40K \newline
6. Unknown \newline

**Gender**\newline
Gender in *BankChurners* was collected in two different categories: male and female. This dataset contains 5358 females and 4769 males. The number of female respondents was more than 11% of the number of male respondents.\newline
\newline
**Credit_Limit, Customer_Age, Total_Trans_Ct** \newline
Credit limit, customer's age and the frequency of the transaction were collected as numeric variables.\newline
\newline

Following are the graphs to show the frequency between the total number of transactions and each income category in different genders. The number of transactions is grouped under 6 intervals.\newline

```{r,echo=FALSE,include=FALSE}
# Input "BankChurners"
databc<-read.csv("/Users/yeonkiii/Documents/uoft/STA303/Prof dev/BankChurners.csv")

library(tidyverse)
```

```{r,echo=FALSE}
# Creating new dataset by drawing out the needed variables
CreditLimit <- databc %>% 
  select('Credit_Limit','Customer_Age','Gender','Education_Level',
         'Income_Category','Total_Trans_Ct','Dependent_count')


# Separate Trans Ct into 6 intervals
ct6 <- CreditLimit %>% 
  mutate(ct_int = case_when(
    Total_Trans_Ct <= 30 ~ 30,
    Total_Trans_Ct <= 51 ~ 51,
    Total_Trans_Ct <= 72 ~ 72,
    Total_Trans_Ct <= 93 ~ 93,
    Total_Trans_Ct <= 114 ~ 114,
    Total_Trans_Ct <= 139 ~ 139
  ))

ct6 <- ct6 %>% 
  mutate(cred_limit = case_when(
    Credit_Limit <= 6951 ~ 2757,
    Credit_Limit <= 12461 ~ 9706,
    Credit_Limit <= 17977 ~ 15219,
    Credit_Limit <= 23490 ~ 20734,
    Credit_Limit <= 29003 ~ 26247,
    Credit_Limit <= 34516 ~ 31760
  ))

# Group out male respondents
ctm <- ct6 %>% 
  filter(Gender == 'M') %>% 
  mutate(Income_Category = fct_relevel(Income_Category, "$40K - $60K", after = 3)) %>% 
  mutate(Income_Category = fct_relevel(Income_Category, "$80K - $120K", after = 1))
                                       
ctm$ct_int <- as.factor(ctm$ct_in)
  
# Group out female respondents
ctf <- ct6 %>% 
  filter(Gender == 'F') 
  

ctf$ct_int <- as.factor(ctf$ct_in)

# Plot the bar graph of total transaction count in different genders
# male

ctm %>% 
  ggplot(aes(x = ct_int, fill= Income_Category)) + 
  geom_bar(position = "stack") + 
  coord_cartesian(ylim=c(0,2000)) + 
  xlab("Interval of Number of Total Transactions") +
  labs(caption = "Figure 1a. Total Number of Transaction (Male)") +
  theme_minimal() +
  scale_fill_manual(values = c("$120K +" = "#F8766D",
                               "$80K - $120K" = "#BB9D00",
                               "$60K - $80K" = "#00B81F",
                               "$40K - $60K" = "#00BFC4",
                               "Less than $40K" = "#529EFF",
                               "Unknown" = "#F066EA"))
# female
ctf %>% 
  ggplot(aes(x = ct_int, fill= Income_Category)) + 
  geom_bar(position = "stack") +
  xlab("Interval of Number of Total Transactions") +
  labs(caption = "Figure 1b. Total Number of Transaction (Female)") +
  theme_minimal() +
  scale_fill_manual(values = c("$40K - $60K" = "#00BFC4",
                               "Less than $40K" = "#529EFF",
                               "Unknown" = "#F066EA"))
  

```
"30": The number of total transaction is between 1 to 30;\newline
"51": The number of total transaction is between 31 to 51;\newline
"72": The number of total transaction is between 52 to 72;\newline
"93": The number of total transaction is between 71 to 93;\newline
"114": The number of total transaction is between 94 to 114;\newline
"139": The number of total transaction is between 115 to 139.\newline

## 2.2 Model
The amount of credit limit was initially predicted based on age, gender, education level, income, and the count of the transactions (which would be referred to as the initial model). These covariates were first introduced to the model because they were commonly collected from banks' customers. In other words, these covariates were the basic information that credit card companies would use to evaluate applicants.\newline

Since the response variable (credit limit) is a continuous variable, and quantitative and qualitative variables were included as explanatory variables, a multivariables linear model would naturally be introduced to the data.\newline

Based on Figure 1a and 1b, the distribution of count of the transaction is different in genders. Thus the count of transaction-by-gender interaction was considered to be added to the linear model.\newline


The initial model is:\newline
$$Credit Limit = \beta_0 + \beta_1X_{age} + \beta_2X_{tc} + \beta_{3}X_{m} + \beta_{4,j}X_{edu,j} + \beta_{5,k}X_{inc,k} +\beta_{6}X_{tc}X_{m} +\epsilon_i $$
where $Credit$ $Limit$ represents the credit limit that respondent has.\newline

$X_{age}$ represents the age of the respondent.\newline

$X_{tc}$ represents the count of the transactions.\newline

$X_{m}$ represents male customer.\newline
$X_{m}=1$ if the respondent is male, and 0 otherwise.\newline

$X_{edu,j}$ represents the education level of the respondent. The subscript *j* means the corresponding *jth* education level from Doctorate, Graduate, High School, Post-Graduate, Uneducated, Unknown, and College.\newline

$X_{inc,k}$ denotes the income category of the respondent. The subscript *k* implies the actual category that the respondent's income falls into. The income categories includes \$120K+, \$80K - \$120K, \$60K - \$80K,  \$40K - \$60K, less than \$40K, and unknown.\newline

$\epsilon_i$ is the independent error term.\newline
\newline


$\beta_0$ shows the expected credit limit of the respondent who is a 0-year-old female, but graduated from college, having income more than \$120K and did not use her credit card, which is unrealistic. Thus $\beta_0$ does not have an intuitive interpretation. \newline

$\beta_1$ is the coefficient of the variable *age*, it shows the effect of age on the credit limit. A negative $\beta_1$ implies that as age increases, the amount of credit limits would decrease by controlling other variables unchanged. Conversely, a positive $\beta_1$ indicates that age is directly proportional to the credit limit.  \newline

$\beta_2$ is the coefficient of the number of transactions, it represents the relationship between the count of the transaction and the amount of the credit limit. In other words, a positive $\beta_2$ implies that there is a positive trend between the number of the transaction and the amount of the credit limit. Conversely. negative $\beta_2$ indicates the negative trend between them. \newline

$\beta_{3}$ shows the difference in the amount of credit limit between male and female respondents. Negative $\beta_{3}$ means the male respondent has a lower credit limit. Conversely, a positive $\beta_{3}$ value indicating a higher credit limit that male respondent has than female respondent. \newline

$\beta_{4,j}$ implies the difference in the amount of credit limit between respondent who has the corresponded(*j*th category in education level) education certificates and the respondent who graduated from college. Similarly, a positive $\beta_{4,j}$ shows that people who have the associated education certificates tend to have higher credit limits than people who graduated from the reference education level of college. Conversely, negative $\beta_{4,j}$ implies that people who graduated from college tend to have higher credit limits. \newline

$\beta_{5,k}$ denoted as the difference in the amount of credit limit between people who has income in the *k*th interval and the reference income of above \$120K. Negative $\beta_{5,k}$ implies that people who have income above \$120K would have higher credit limits than those who have income in the associated interval. Conversely, positive $\beta_{5,k}$ shows that people who have income in *k*th interval tend to have higher credit limits than those who have income above $120K. \newline

$\beta_{6}$ measures the change in size of the effect of the number of total transactions on the amount of credit limit due to the gender of customers. Negative $\beta_{6}$ means a negative effect of the total number of transaction on the credit limit that male has. Positive $\beta_{6}$ indicating males' transaction counts have a positive effect on the amount of credit limit. \newline


```{r, echo=FALSE}

# Creating initial model
model0 <- lm(Credit_Limit~Customer_Age + Total_Trans_Ct +
                  Gender + Education_Level +
                  Income_Category  +
                  Total_Trans_Ct * Gender , 
                data = CreditLimit)
#summary(model0)
knitr::kable(broom::tidy(model0), caption = "Initial Model")

```

In order to determine the useful explanatory variables out of 6 mentioned ones, the AIC backward elimination method was applied.\newline

The idea of the AIC backward elimination method is that starting with the model contains all predictors, the predictor that has the highest AIC value would be removed by using *R*. Then refit the model and continue to remove the variable that has the highest AIC value until the model reaches the lowest AIC value. The latest model would be considered as the final model. \newline

```{r, echo=FALSE, results=FALSE}
model1<-step(model0, direction = "backward")
```

```{r, echo=FALSE}
#summary(model1)
knitr::kable(broom::tidy(model1), caption = "Reduced Model")
```

Based on the result of AIC elimination, age and education level were removed from the model, and the final model is as following:\newline
$$CreditLimit = \beta_{0} + \beta_1X_{tc} + \beta_{2}X_{m} + \beta_{3,k}X_{inc,k} +\beta_{4}X_{tc}X_{m} +\epsilon_i $$\newline
$Credit$ $Limit$ represents the credit limit that respondent has.\newline

$X_{tc}$ represents the count of total transactions.\newline

$X_{m}$ represents male customer.\newline
$X_{m}=1$ if the respondent is male, and 0, otherwise.\newline

$X_{inc,k}$ denotes the income category of the respondent. The subscript *k* implies the actual category that the respondent's income falls into. The income categories includes \$120K+, \$80K - \$120K, \$60K - \$80K, \$40K - \$60K, less than \$40K, and unknown.\newline

$\epsilon_i$ is the independent error term.\newline
\newline
$\beta_0$ shows the credit limit of the female respondent who graduated from college, having income more than \$120K and did not use her credit card at all. \newline

$\beta_{1}$ is the coefficient of the number of transactions, it represents the relationship between the count of the transaction and the amount of the credit limit. In other words, a positive $\beta_1$ implies that there is a positive trend between the number of the transaction and the amount of the credit limit. Conversely. negative $\beta_1$ indicates the negative trend between them. \newline

$\beta_{2}$ represents the difference in the amount of credit limit between male and the reference gender of female. Negative $\beta_{2}$ means the male respondent has a lower credit limit. Conversely, positive $\beta_{2}$ value indicating a higher credit limit that male respondent has than female respondent. \newline

$\beta_{3,k}$ denoted as the difference in the amount of credit limit between people who has income in the *k*th interval and the reference income of above \$120K. 

$\beta_{4}$ measures the change in size of the effect of the number of total transactions on the amount of credit limit due to the gender of customers. Negative $\beta_{4}$ means a negative effect of the total number of transaction on the credit limit that male has. Positive $\beta_{4}$ indicating males' transaction counts has a positive effect on the amount of credit limit. \newline

## 3. Result
Figure 1a and 1b shows different distributions of total transaction count between males and females, thus it is reasonable that gender has difference effect on count of transaction, which implies that interaction term should be included in the model.

Based on Table 1, p-value for education levels and age were greater than 5% significant level, thus all the education levels and age were not significant associated with the amount of credit limits. \newline

The AIC backward method was applied to determine the useful predictors out of the 6 variables. Based on table 2, we could see that the p-value for gender, income, and count of transaction were all less than 5% significant level. It means that among all the mentioned covariates, only respondents' gender, income, and count of transaction were associated with the amount of credit limits.\newline

According to Table 2, $\beta_{1}$ is 19.029, which means that controlling other variables, as the count of transaction increases, the amount of credit limit would also increase.\newline

$\beta_{2}$ is -1110.711 shows that controlling other variables unchanged, male respondents have credit limit $1110.71 lower than female respondents.\newline

Negative $\beta_{3,k}$ indicates that fixing other variables, people who have higher income would have higher credit limits as well. But surprisingly, people who did not provide income information would still have higher credit limit than whose income between \$60K - \$80K.\newline

# 4. Limitations

Due to the privacy issue, some information of the data was not revealed, which would raise some red flags. The observation units were only labeled as 9-digit numbers, we were not able to make sure whether the observation units were from the same household or not. If they were, then the units were not independent. For example, if one family member(other than the household head) were purchasing all the daily supplies, then she/he would have more transactions than other members; the household head could possibly have less transaction but higher income.

Moreover, as mentioned before, the data were lack of background information such as the collecting method, which could lower the accuracy of the analysis.


# 5. Conclusion
The prediction of credit limits was initially made based on the basic covariates: age, gender, education level, income, frequency of transaction, and transaction count-by-gender interaction. The AIC method suggested that only gender, income, frequency of transactions, and transaction count-by-gender interaction were associated with credit limits. Females tended to have a higher credit limit than males under the same conditions of other covariates. Fixing other variables unchanged, people with higher income were more possible to have higher credit limits. Moreover, under the same conditions of other covariates, as the frequency of transaction increases, credit limits would also increase. \newline

There are still some aspects that require further research on it, such as the background information of the dataset, and the identity of the observation units.\newline


\newpage
## Reference
Goyal, S. (2020, November 19). Credit Card customers. Retrieved April 2, 2021, from https://www.kaggle.com/sakshigoyal7/credit-card-customers

Porter, W., Staff, C., McClanahan, A., DeNicola, L., Hipp, D., &amp; Proctor, C. (2020, November 16). What Is a Credit Limit and How Is It Determined? Retrieved April 2, 2021, from https://www.creditkarma.com/advice/i/what-is-a-credit-limit

Extract default color palette of ggplot2 r package (example): Hex codes. (2020, September 30). Retrieved April 2, 2021, from https://statisticsglobe.com/identify-default-color-palette-names-of-ggplot2-in-r

N/A, N. (n.d.). CRITERION-BASED PROCEDURES. Retrieved April 2, 2021, from http://www.biostat.jhsph.edu/~iruczins/teaching/jf/ch10.pdf



